Global patterns of forest autotrophic carbon fluxes

### 2020-01-19 manuscript/manuscript draft.Rmd
working on Helene's comments
finished through results.

Models were run for total annual FACF against annual climate variables, and for monthly growing season fluxes, defined as annual FACF/length of growing season, against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.3). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect.

For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone. However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation. Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity. GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (fig. 4). ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it. Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on productivity.

We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig. 4). Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.

*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity. For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes. There was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05). Productivity increased with solar radiation for GPP ($R^2$ = 0.21, p<0.001), NPP ($R^2$ = 0.21, p<0.001), BNPP ($R^2$ = 0.16, p<0.001) and BNPP~fine~ ~root~ ($R^2$ = 0.12, p<0.01), and with PET for GPP ($R^2$ = 0.15, p<0.01), NPP ($R^2$ = 0.18, p<0.01), BNPP ($R^2$ = 0.23, p<0.0001), BNPP~fine~ ~root~ ($R^2$ = 0.11, p<0.05), and ANPP~woody~ ~stem~ ($R^2$ = 0.06, p<0.05).

>>>>

Models were run for total annual FACF against annual climate variables, and for fluxes standardized by growing season length, defined as annual FACF/length of growing season, against growing season climate variables. For analyses on data within the growing season, only linear models were specified.

@@

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.3). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect.  {**NEED SI TABLE GIVING FITTED COEFFICIENTS & STATS**}

For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone {**Helene: but less than MATxMAP? if so, why bother mentioning?**}. However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.

```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and mean annual precipitation on FACF. For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*How does productivity relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity. In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.  {**cite some table/ SI?**}

@@

*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (Fig. 4) {**give statistic on signficiance**}. ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality. Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}. Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. In contrast, there was no significant effect of precipitation seasonality on FACF.

We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase in productivity with length of growing season (Fig. 4). Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.

*Within the growing season, how does FACF vary with climate?* When FACF was standardized by growing season length, correlations with climate variables were much weaker. Temperature, precipitation, PET, and solar radiation each had a small effect on one or more FACF (Fig. **S#**). Specifically, there was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05). Productivity increased with solar radiation for GPP ($R^2$ = 0.21, p<0.001), NPP ($R^2$ = 0.21, p<0.001), BNPP ($R^2$ = 0.16, p<0.001) and BNPP~fine~ ~root~ ($R^2$ = 0.12, p<0.01), and with PET for GPP ($R^2$ = 0.15, p<0.01), NPP ($R^2$ = 0.18, p<0.01), BNPP ($R^2$ = 0.23, p<0.0001), BNPP~fine~ ~root~ ($R^2$ = 0.11, p<0.05), and ANPP~woody~ ~stem~ ($R^2$ = 0.06, p<0.05).  {**site SI figure, stats might work better in a table**}  {**did we look at respiration in this anlaysis?**}
